Paper 1

### 2022-06-14 scripts/Paper 1.Rmd
update

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(plotly)
library(plyr)
library(dplyr)
library(knitr)
library(tidyverse)
#library(devtools)
library(openair)
library(tidyr)
library(naniar)
library(forecast)
library(kableExtra)
library(htmltools)
library(ggpubr)
library(rstatix)
library(broom)

@@

```{r create daily mean dataframe, include=FALSE}
# add variables to this daily df as needed
input.daily <- input %>%
  drop_na(Timestamp) %>% # some NAs in the Timestamp for some reason. Need to look into
  ddply(c("Year", "jday", "month"), summarise,
        Precip = sum(Precip, na.rm = TRUE),
        Ta.mean = mean(Ta, na.rm = TRUE),
        Ta.max = max(Ta, na.rm = TRUE),
        Ta.min = min(Ta, na.rm = TRUE),
        Ts.mean = mean(Ts, na.rm = TRUE),
        Ts.max = max(Ts, na.rm = TRUE),
        Ts.min = min(Ts, na.rm = TRUE),
        SWC = mean(SWC, na.rm = TRUE),
        RH = mean(RH, na.rm = TRUE),
        VPD = mean(VPD, na.rm = TRUE),
        NEE_LT = mean(NEE_LT, na.rm = TRUE), #daily average umol m-2 s-1
        ER_LT = mean(ER_LT, na.rm = TRUE),
        GPP_LT = mean(GPP_LT, na.rm = TRUE)) %>%
  mutate(NEE_LT.gC = gofC.f(NEE_LT),  #converts to daily total NEE in g C CO2
         ER_LT.gC = gofC.f(ER_LT),
         GPP_LT.gC = gofC.f(GPP_LT))
```

@@

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 10, fig.height = 4}
input.daily %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-1]
colnames(rainfall_L)[4] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[3] <- "jday"
colnames(rainfall_L)[2] <- "month"
rainfall_L$Site <- "L"

@@

rainfall <- rbind(rainfall_L, rainfall.SP)
rainfall$month <- as.factor(rainfall$month) 
rainfall$month <- month.abb[rainfall$month]

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()

@@

```{r Air temperature, echo=FALSE, message = FALSE, fig.width = 10, fig.height = 4}
ggplotly(ggplot(input.daily, aes(x = jday, y = Ta.mean, group = Year)) +
  geom_line(aes(color = Year))+
     geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
  theme_classic())

@@

```{r, echo=FALSE, message = FALSE, fig.width = 10, fig.height = 4}
ggplotly(ggplot(input.daily, aes(jday, SWC, group = Year)) +
  geom_line(aes(color = Year))+
  xlab('Julian day of year') +
  ylab('SWC %') +
  theme_classic())
```

@@

```{r timeseries analysis}
tsData <- EuStockMarkets[, 1] # ts data
decomposedRes <- decompose(tsData, type="mult") # use type = "additive" for additive components
plot (decomposedRes) # see plot below
stlRes <- stl(tsData, s.window = "periodic")

input.daily %>% anova_test(NEE_LT.gC~Precip+season)

rainfall$TS <- as.Date(with(rainfall, paste(Year, month, jday,sep="-")), "%Y-%m-%d")

TSData <- rainfall[, 1] # ts data
decomposedRes <- decompose(tsData, type="mult") # use type = "additive" for additive components
plot (decomposedRes) # see plot below
stlRes <- stl(tsData, s.window = "periodic")
```

>>>>

```{r setup, include=FALSE}
library(data.table)
library(janitor)
library(plotly)
library(plyr)
library(dplyr)
library(knitr)
library(tidyverse)
#library(devtools)
library(openair)
library(tidyr)
library(naniar)
library(forecast)
library(kableExtra)
library(htmltools)
library(ggpubr)
library(rstatix)
library(broom)
library(gtable)
library(gridExtra)
library(cowplot)

@@

```{r create daily mean dataframe, include=FALSE}
# add variables to this daily df as needed
input.daily <- input %>%
  drop_na(Timestamp) %>% # some NAs in the Timestamp for some reason. Need to look into
  ddply(c("Year", "jday", "month"), summarise,
        Precip = sum(Precip, na.rm = TRUE),
        Ta.mean = mean(Ta, na.rm = TRUE),
        Ta.max = max(Ta, na.rm = TRUE),
        Ta.min = min(Ta, na.rm = TRUE),
        Ts.mean = mean(Ts, na.rm = TRUE),
        Ts.max = max(Ts, na.rm = TRUE),
        Ts.min = min(Ts, na.rm = TRUE),
        SWC = mean(SWC, na.rm = TRUE),
        RH = mean(RH, na.rm = TRUE),
        VPD = mean(VPD, na.rm = TRUE),
        NEE_LT = mean(NEE_LT, na.rm = TRUE), #daily average umol m-2 s-1
        ER_LT = mean(ER_LT, na.rm = TRUE),
        GPP_LT = mean(GPP_LT, na.rm = TRUE)) %>%
  mutate(NEE_LT.gC = gofC.f(NEE_LT),  #converts to daily total NEE in g C CO2
         ER_LT.gC = gofC.f(ER_LT),
         GPP_LT.gC = gofC.f(GPP_LT))

input.daily$month <- as.numeric(input.daily$month) 
input.daily$month <- month.abb[input.daily$month]

@@

```{r Precipitation, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}

(SP.rain <- input.daily %>% mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       #scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
   scale_x_discrete(limits = month.abb) +
  theme_classic()+
  theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

#below is precip data from liaweenee - it only goes up to Jan 2022 right now
rainfall_L <- read.csv("R:/SET/PlantSci/Shared Work Spaces/Ecology/Peoples files/Marion Nyberg/PhD/Rainfall/rainfall.csv")
rainfall_L <- rainfall_L[-4]
colnames(rainfall_L)[5] <- "Precip"
rainfall_L$Year <- as.factor(rainfall_L$Year)
rainfall_L <- rainfall_L %>% filter(Year == 2019)
colnames(rainfall_L)[1] <- "jday"
colnames(rainfall_L)[3] <- "month"
rainfall_L$Site <- "L"
rainfall_L <- rainfall_L %>% filter(month >= 9)

@@

rainfall <- rbind(rainfall_L, rainfall.SP)
rainfall$month <- as.numeric(rainfall$month) 
rainfall$month.abb <- month.abb[rainfall$month]

rainfall %>%  mutate(Precip = replace_na(Precip, 0)) %>%
  ddply(c("Year", "month"), summarise,
  monthly.total = sum(Precip)) %>%
  ggplot(., aes(x = month, y = monthly.total, group = Year, fill = Year)) +
       scale_x_discrete(limits = month.abb) +
  geom_col(position = "dodge") + 
  ylab("Monthly rainfall (mm)")+
  xlab("Month")+
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))

@@

```{r Air temperature, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
TA <- ggplot(input.daily, aes(x = jday, y = Ta.mean, group = Year)) +
  geom_line(aes(color = Year), size =1)+
    xlab("Julian day of year")+
    ylab("Air temperature (°C)")+
     #geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))

(TS <- ggplot(input.daily, aes(x = jday, y = Ts.mean, group = Year)) +
  geom_line(aes(color = Year), size =1)+
    xlab("Julian day of year")+
    ylab("Soil temperature (°C)")+
     #geom_ribbon(aes(ymin=Ta.min, ymax=Ta.max), alpha=.3, linetype=0) +
  theme_classic()+
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14)))

@@

```{r, echo=FALSE, message = FALSE, fig.width = 8, fig.height = 4}
SWC <- ggplot(input.daily, aes(jday, SWC, group = Year)) +
  geom_line(aes(color = Year), size = 1.5)+
  xlab('Julian day of year') +
  ylab('SWC %') +
  theme_classic() + 
    theme(axis.title = element_text(size = 14),
          axis.text = element_text(size = 13),
          legend.text = element_text(size = 13),
          legend.title = element_text(size = 14))
```

```{r, fig.width = 10, fig.height = 6}
Legend.met<- get_legend(SP.rain+ theme(legend.position = 'right'))

Fig1 <- plot_grid(SP.rain + theme(legend.position="none"), TA+ theme(legend.position="none"),
             SWC+ theme(legend.position="none"), TS+ theme(legend.position="none"), ncol = 2, rel_heights = c(0.5, 0.5, 0.5, 0.5))

Fig1
tiff("Fig1.tiff", units="in", width=10, height=6,  res=300)
print(Fig1)
dev.off()

Legend.met
```

@@
